---
title: "ClusterMap"
author: "Daniel Russo"
date: "1/26/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reticulate)
use_condaenv(condaenv = 'cheminf')

```

## Set up

The variable `NAMES` can be set to `AOP` or `MECHANISM`.  `AOP` means set the 
variable to the AOP name like AOP 1, AOP 2, etc.  `MECHANISM` means set it to the
mechanism like CYTOTOXICITY, etc.

## data clean up

use python to set all aops bound all AOPs < 0 to 0 AUC scores can sometimes result in 
less than 0 but this is meaningless.

NOTE: for whatever reason, the `toxcast2` environment causes the R kernel to crash.  Hence, I am using the `cheminf` environment

```{python}
import pandas as pd

NAMES = "KE"

df = pd.read_csv('../data/text/full_aop_scores.csv')
clusters = pd.read_csv('../data/text/full_tox21_clusters.csv')
names_frame = pd.read_csv("../data/text/AOP_mechanism_table.csv")
if NAMES == "AOP":
    mapper = dict(zip(names_frame["Mechanism"], names_frame["AOP"]))
    df['aop'] = df['aop'].map(mapper)

if NAMES == "KE":
    mapper = dict(zip(names_frame["Mechanism"], names_frame["KE"]))
    df['KE'] = df['aop'].map(mapper)

df.loc[df.auc_score <= 0, 'auc_score'] = 0

df_wide = df.pivot(index='casn', columns='KE', values='auc_score')


df_wide.head()

```

### Calculate KE scores by cluster

Group by cluster and take the mean per cluster for AOP score

Remove columns that have no AOP scores > 0


```{python}

cluster_aops = df.merge(clusters[['casn', 'Cluster']]).groupby(['Cluster', 'KE'])['auc_score'].mean()

cluster_aops = cluster_aops.reset_index().pivot(index='Cluster', values='auc_score', columns='KE')

# remove AOPs that are 0
# for every cluster to make 
# the visualization better
cluster_aops = cluster_aops.loc[:, (cluster_aops  > 0).any()]

```

Arbitrarily set the min as 0, 0.1 as the midpoint and finally the max as red

```{r, echo=FALSE}
library(ComplexHeatmap)
library(circlize)

low.color = scales::muted("blue")
#low.color = "blue"
mid.color = "white"
high.color = scales::muted("red")
#high.color = "red"

cluster_means = py$cluster_aops

# can change this to max(cluster_means)
# to replace the 1, to make the differences 
# more apparent
col.mapper = colorRamp2(c(min(cluster_means), 0.1, 1), c(low.color, mid.color, high.color))

#png("../data/figures/cluster_heatmap.png", width=9, height=8,units="in",res=800)
Heatmap(as.matrix(cluster_means ),
        col=col.mapper,
        rect_gp = gpar(col = "grey", lwd = 1),
        name = "KE Score",
        )
#dev.off()

```


### now plot the interesting Clusters

First prepare the data in pandas


```{python}

good_clusters = [377, 380, 387, 176, 16, 196, 26, 231]
print(clusters)
clusters_limited = clusters[clusters.Cluster.isin(good_clusters)]

limited_df = df.merge(clusters_limited[['casn', 'Cluster', 'Liver',  'Hepatobiliary', 'Hepatocellular']])

limited_df = limited_df.pivot(index=['casn', 'Cluster', 'Liver',  'Hepatobiliary', 'Hepatocellular'], values='auc_score', columns='KE').reset_index()

limited_clusters = limited_df.pop('Cluster').values.tolist()
limited_liver = limited_df.pop('Liver').values.tolist()
limited_hb = limited_df.pop('Hepatobiliary').values.tolist()
limited_hc = limited_df.pop('Hepatocellular').values.tolist()

limited_df = limited_df.set_index('casn')
print(limited_df.mean())

limited_df = limited_df.loc[:, (limited_df  > 0 ).any()]

```

```{r, dev='png'}

aop_scores = py$limited_df
split = py$limited_clusters
liverLimit = py$limited_liver
hcLimit = py$limited_hc
hbLimit = py$limited_hb
print(split)

hbLimit[is.nan(hbLimit)]<-NA
hcLimit[is.nan(hcLimit)]<-NA
liverLimit[is.nan(liverLimit)]<-NA

print(typeof(liverLimit[4]))

#col_fun1 = colorRamp2(c(0, max(aop_scores)), c("white", "red"))
col.mapper = colorRamp2(c(0, 0.1, 1), c(low.color, mid.color, high.color))
liver.col.mapper = colorRamp2(c(0, 1), c("green", "red"))

ha = rowAnnotation(
  Hepatotoxicity = factor(liverLimit),
  Hepatocellular = factor(hcLimit),
  Hepatobiliary = factor(hbLimit),
  col=list(Hepatotoxicity=c("1" = "red", "0" = "Green"),
           Hepatobiliary=c("1" = "red", "0" = "Green"),
           Hepatocellular=c("1" = "red", "0" = "Green")),
  name="Liver",
  gp = gpar(col = "grey", lwd = 1)
)

#png("../data/figures/full_heatmap.png", width=9, height=8,units="in",res=800)
Heatmap(as.matrix(aop_scores),
        split = factor(paste("", split)),
        col=col.mapper,
        rect_gp = gpar(col = "grey", lwd = 1),
        name = "KE Score", border_gp = gpar(col = "black", lty = 2),
        left_annotation = ha,
        row_labels = rep("", nrow(aop_scores)))
#dev.off()
```

